### 2015-09-23 envolve.tex
reconstruct logically

% !TEX root = Eco-Model.tex
\section{Manufacturing Ecosystem Envolving} % (fold)
\label{sec:envolve}
the factors of evolving/ howto evolving

\subsection{Simulation} % (fold)
\label{sub:simulation}

% subsection simulation (end)

\subsection{Solving} % (fold)
\label{sub:solving}

% subsection solving (end)

\subsection{Control} % (fold)
\label{sub:control}

% subsection control (end)
% section envolve (end)