### 2015-09-27 envolve.tex
init on mac

% !TEX root = Eco-Model.tex
\section{Manufacturing Ecosystem Envolving} % (fold)
\label{sec:envolve}
The factors of evolving/ howto evolving

\subsection{Envolving Aim} % (fold)
\label{sub:envolving_aim}

% subsection envolving_aim (end)

\subsection{Main Factors} % (fold)
\label{sub:main_factors}

% subsection main_factors (end)

\subsection{Control of The Chain} % (fold)
\label{sub:control}

% subsection control (end)
% section envolve (end)