### 2015-10-02 envolve.tex
give up simulation

% !TEX root = Eco-Model.tex
\section{Manufacturing Ecosystem Envolving} % (fold)
\label{sec:envolve}
The factors of evolving/ howto evolving

\subsection{Envolving Aim} % (fold)
\label{sub:envolving_aim}

% subsection envolving_aim (end)
\subsection{Main Factors} % (fold)
\label{sub:main_factors}

% subsection main_factors (end)

\subsection{Rating Schema} % (fold)
\label{sub:rating_schema}

% subsection rating_schema (end)

\subsection{Recommand System} % (fold)
\label{sub:recommand_system}

% subsection recommand_system (end)

\subsection{Control of The Chain} % (fold)
\label{sub:control}

% subsection control (end)
% section envolve (end)