### 2015-10-02 schedule.tex

% !TEX root = Eco-Model.tex
\section{Services Scheduling} % (fold)
\label{sec:schedule}

In order to properly accomplish the proposed tasks with optimal allocation of the manufacturing resources, a series of methodologies were combined to assist the manufacturing services scheduling process in Cloud Manufacturing Ecosystem, since manufacturing serivices are contrusted from manufacturing resources.

OAS

Most of the related methodologies can be archived into \textbf{four} main modules, namely \textit{Decision Making Support}, \textit{Rating Schema}, \textit{Recommand System} and \textit{Matching Regulation}, which are going to explained in detail respectively.

The scheduling effect can be estimated and observed through \textit{Service Dashboard}, which also provides the control function that will help the evolution in the Ecosystem.

\subsection{Decision Making Support} % (fold)
\label{sub:decision_making_support}

% subsection decision_making_support (end)

\subsection{Matching Regulation} % (fold)
\label{sub:matching_regulation}

% subsection matching_regulation (end)

\subsection{Service Dashboard} % (fold)
\label{sub:service_dashboard}

% subsection service_dashboard (end)
% section model (end)