### 2015-10-02 figures/platformstruct.tex
add some node to the platform structure figure

% !TEX root = ..\Eco-Model.tex
\begin{tikzpicture}
[mindmap,
every node/.style={concept, text=white, execute at begin node=\hskip0pt,circular drop shadow},
root concept/.append style={concept color=black, fill=white, line width=1ex, text=black, font=\Large\scshape},
grow cyclic,
level 1/.append style={level distance=4.5cm,sibling angle=90, font=\scshape},
level 2/.append style={level distance=3cm,sibling angle=45, font=\scriptsize\bf},
level 3/.append style={level distance=2cm,sibling angle=60, font=\scriptsize\bf},
level 4/.append style={level distance=1cm,sibling angle=60, font=\tiny\bf}]
\node [root concept] {Cloud Manufacturing Ecosystem}  % root
child [concept color = gray]{ node {Schedule}
child { node {Service Standardization}}
child { node {Decision Making Support}}
child { node {Platform Assistance}}
child { node {Match Regualtion}}
}
child [concept color = gray]{ node {User}
  child { node {Submit Order}}
  child { node {Provide Resource}}
  child { node {Make Transaction}}
  child { node {Evalute Others}}
  }
child [concept color = gray]{ node {Envolve}
  child { node {Optimization}
    child { node {Manufacturing}}
    child { node {Industrial Chain}}
    }
  child { node {Member}
    child { node {Introduce}}
    child { node {Eliminate}}
    }
  child { node {Adjust User Proportion}
    child { node {Rating}}
    child { node {Recommand}}
    }
  child { node {Development Guidance}}
  }
child [concept color = gray]{ node {Platform}
  child { node {Authority Control}}
  child { node {Transaction Bridge}}
  child { node {Dashbord}
    child { node {Service}}
    child { node {Task}}
    child { node {User}}
    }
child { node {Administrate}}
  child { node {Hub}
    child { node {Service}}
    child { node {Resource}}
    child { node {Ranking}}
    }
  child { node {Scheduling Assistance}}
}
;
\end{tikzpicture}