### 2015-10-02 schedule.tex

% !TEX root = Eco-Model.tex
\section{Services Scheduling} % (fold)
\label{sec:schedule}

In order to properly accomplish the proposed tasks with optimal allocation of the manufacturing resources, a series of methodologies were combined to assist the manufacturing services scheduling process in Cloud Manufacturing Ecosystem, since manufacturing serivices are contrusted from manufacturing resources.

OAS  
\begin{figure}[!h]
\centering\small\resizebox{0.6\textwidth}{!}{\input{figures/order_flow}}
\caption{Primary Flow of Order}
\label{fig:orderflow}
\end{figure}
Most of the related methodologies can be archived into \textbf{four} main modules, namely \textit{Service Standardization}, \textit{Decision Making Support}, \textit{Matching Regulation} and \textit{Platform Assistance} which are going to explained in detail respectively.

The scheduling effect can be estimated and observed through \textit{Service Dashboard}, which also provides the control function that will help the evolution in the Ecosystem.

\subsection{Service Standardization} % (fold)
\label{sub:service_standardization}

% subsection service_standardization (end)

\subsection{Decision Making Support} % (fold)
\label{sub:decision_making_support}

% subsection decision_making_support (end)

\subsection{Matching Regulation} % (fold)
\label{sub:matching_regulation}

% subsection matching_regulation (end)

\subsection{Platform Assistance} % (fold)
\label{sub:platform_assistance}

% subsection platform_assistance (end)

\subsection{Service Dashboard} % (fold)
\label{sub:service_dashboard}

% subsection service_dashboard (end)
% section model (end)