### 2015-10-07 beamer/envolve.tex
main struct

% !TEX root = ../beamer.tex
\section{Manufacturing Ecosystem Envolving}
\subsection{Envolving Aim \& Main Factors}
\begin{frame}{Manufacturing Ecosystem Envolving}{Envolving Aim \& Main Factors}

\end{frame}

\subsection{Rating Schema}
\begin{frame}{Manufacturing Ecosystem Envolving}{Rating Schema}

\end{frame}

\subsection{Recommand System}
\begin{frame}{Manufacturing Ecosystem Envolving}{Recommand System}

\end{frame}

\subsection{User Introduction \& Elimination}
\begin{frame}{Manufacturing Ecosystem Envolving}{User Introduction \& Elimination}

\end{frame}