### 2015-10-07 beamer/intro.tex

% !TEX root = ../beamer.tex
\section{Introduction and Preliminary}

\subsection{Our Projects}

\begin{frame}{Introduction}{Our Projects}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\subsection{Operating Mode}

% You can reveal the parts of a slide one at a time
% with the \pause command:
\begin{frame}{Preliminary}{Operating Mode}
  \begin{itemize}
  \item {
    First item.
    \pause % The slide will pause after showing the first item
  }
  \item {   
    Second item.
  }
  % You can also specify when the content should appear
  % by using <n->:
  \item<3-> {
    Third item.
  }
  \item<4-> {
    Fourth item.
  }
  % or you can use the \uncover command to reveal general
  % content (not just \items):
  \item<5-> {
    Fifth item. \uncover<6->{Extra text in the fifth item.}
  }
  \end{itemize}
\end{frame}