### 2015-10-07 beamer/schedule.tex

% !TEX root = ../beamer.tex
\section{Services Scheduling}
\subsection{Standardization}
\begin{frame}{Services Scheduling}{Main}

\end{frame}
\subsection{Decision Making Support}
\begin{frame}{Services Scheduling}{Decision Making Support}

\end{frame}

\subsection{Matching Regualtion}
\begin{frame}{Services Scheduling}{Matching Regualtion}

\end{frame}

\subsection{Platform Assistance}
\begin{frame}{Services Scheduling}{Platform Assistance}

\end{frame}