### 2015-10-14 beamer/envolve.tex
init beamer

% !TEX root = ../beamer.tex
\section{Manufacturing Ecosystem Envolving}
\subsection{Envolving Aim \& Main Factors}
\begin{frame}{Envolving Aim \& Main Factors}{Manufacturing Ecosystem Envolving}
\onslide<+->{The aim of envolving in the Ecosystem is to: \begin{itemize}
\item optimize the utilize of manufacturing resources
\item find the adaptive operating model for enterprises
\item filter out the underisable enterprises.
\end{itemize}}
\onslide<+->{
\begin{block}{Main Factors for Envolving}
\begin{itemize}
\item Rating Schema;
\item Recommand System and
\item Criterion of User Introduction and Elimination
\end{itemize}
\end{block}}
\end{frame}

\subsection{Rating Schema}
\begin{frame}{Rating Schema}{Manufacturing Ecosystem Envolving}
\onslide<+->{Rating user's rank is the base of Recommand System, so the rating schema should be designed. Most modern rating systems are derived from Elo's model:}
\onslide<+->{\begin{numcases}{}
E_A = \frac{1}{1+10^{\frac{R_B-R_A}{400}}} \\
R_A := R_A + K(S_A - E_A) \\
E_B = \frac{1}{1+10^{\frac{R_A-R_B}{400}}} \\
R_B := R_B + K(S_B - E_B)
\end{numcases}}
\end{frame}

\subsection{Recommand System}
\begin{frame}{Recommand System}{Manufacturing Ecosystem Envolving}
\onslide<+->{There are two scenarios in recommendation:
\begin{itemize}
\item Service Searching
\item Platform priority setting
\end{itemize}
}
\onslide<+->{
	Collaboration Filtering is popular method used in recommand system:
	\begin{equation}
	\begin{split}
	\min_{p_*,q_*,b_*} & \sum_{(u,i)\in\kappa}\left(r_{ui} -\mu - b_u - b_i - p^T_uq_i \right)^2 \\
		& + \lambda_3 \left( \|p_u\|^2 + \|q_i\|^2 + b_u^2 + b_i^2 \right)
	\end{split}
	\end{equation}
	\begin{numcases}{}
	\hat{r}_{ui} = b_{ui} + p^T_uq_i \\
	b_{ui} = \mu + b_u + b_i  \\
	\kappa =\{(u,i)|r_{ui} \text{is known}\} 
	\end{numcases}
}
\end{frame}

\subsection{User Introduction \& Elimination}
\begin{frame}{User Introduction \& Elimination}{Manufacturing Ecosystem Envolving}
\onslide<+->{
	Ecosystem envolving need a good user introduction and elimination method with:
}
\onslide<+->{
	\begin{block}{Envolving}
	\begin{itemize}
	\item Reinforcement Learning
	\item Control
	\end{itemize}
	\end{block}
}
\end{frame}

\subsection{Envision of the Ecosystem}
\begin{frame}{Envision}{After a long running time and the system seems stable}
\onslide<+->{
	With poper scheduling and envolving, Cloud Manufacturing Ecosystem running with a stable mode:
	\begin{block}{Stable System Envision}
	\begin{itemize}
		\item Less-Popular manufacturing services are utilized instead of idleness with Coupon
		\item Products are manufactured with appropriate resource
		\item High rank users in different type of manufacturing resource lead the industrial chain
		\item Good operating models appear and being refined with time goes by
	\end{itemize}
	\end{block}
}
\end{frame}