### 2015-10-14 beamer/intro.tex

% !TEX root = ../beamer.tex
\section{Introduction}

\subsection{Cloud Manufacturing}
\begin{frame}{Brief Introduction}{Cloud Manufacturing}
\onslide<+->{Cloud manufacturing combines technologies such as
\begin{itemize}
\item Internet of Things
\item Cloud computing
\item Semantic Web
\item Virtualisation
\item Service-oriented technologies
\end{itemize}
}

\onslide<+->{Aims the manufacturing process to be
\begin{itemize}
\item networked
\item intelligent
\item service-oriented
\item knowledge-based
\item energy efficient
\end{itemize}}
\end{frame}

\subsection{Current Items}

\begin{frame}{Current Items}{National High-Tech. R\&D Program, China}
\textbf{Item}: \textit{Key Technologies Research on Cloud Manufacturing Platform for Group Enterprises (No. 2015AA042101)}
\begin{block}{Main Tasks}
  \begin{itemize}
  \item Business model and standard making in group enterprises;
  \item Multi-tenancy technology for on-demand useage;
  \item Platform development for enterprises (Cooperate);
  \item Application demonstration (Cooperate).
  \end{itemize}
\end{block}
\end{frame}

\begin{frame}{Current Items}{National Natural Science Fundation, China}
\textbf{Item}: \textit{BOSS based Product Manufacturing-service Modeling and Cloud Platform Guided Service-network Evolving (No. 71571161)}
\begin{block}{Main Tasks}
\begin{itemize}
\item Methodology study for standard Manufacturing-service;
\item Methodology study for Manufacturing-service evalution;
\item Optimize the matching for service and task;
\item Explore the evolution of Manufacturing-service.
\end{itemize}
\end{block}
\end{frame}