### 2015-11-16 beamer/schedule.tex

% !TEX root = ../beamer.tex
\section{Services Scheduling}
\subsection{Standardization}
\begin{frame}{Support Module}{Service Standardization}
\onslide<+->{The basic support modules for Service Standardization include:
\begin{block}{Basic Modules}
\begin{itemize}
\item Resource Virtualization \& Encapsulation
\item Clustering Method
\item Imporvement Method
\end{itemize}
\end{block}}

\onslide<+->{
Important support Service Scheduling process modules:
\begin{block}{Support Modules}
\begin{itemize}
\item Decision Making Support
\item Matching Regualtion
\item Platform Assistance
\end{itemize}
\end{block}
}
\end{frame}

\subsection{Decision Making Support}
\begin{frame}{Decision Making Support}{Services Scheduling}
\onslide<+->{
	Decisions that users make in Cloud Manufacturing platform include:
}
\onslide<+->{
	\begin{itemize}
	\item Order acceptance;
	\item Task decomposition;
	\item Task (re-)outsourscing;
	\item Seller selection;
	\item Manufacturing arrangement.
	\end{itemize}
}
\end{frame}

\subsection{Matching Regualtion}
\begin{frame}{Matching Regualtion}{Services Scheduling}
\onslide<+->{
	In order to recommand poper manufacturing service with the order user submitted, Cloud Manufacturing platfrom should provide best-match solution, thus service search engine need to adopt a series of well designed matching regulations.
}
\end{frame}

\subsection{Platform Assistance}
\begin{frame}{Platform Assistance}{Services Scheduling}
\onslide<+->{
	Administrator assists platform to set priority for taks or services, so two aspects should be considered for users:
}
\onslide<+->{
	\begin{itemize}
	\item How to make priority setting recommendation
	\item How to make decisions with the recommendation
	\end{itemize}
}
\end{frame}