### 2016-03-04 method.tex
add the enterprise model

% !TEX root = Eco-Model.tex
\section{Methodological Approach} % (fold)
\label{sec:methodological_approach}

\subsection{Modeling} % (fold)
\label{sub:modeling}

\subsubsection{Assumptions and Notation}

Ecosystem application model including:
\subsubsection{Enterprise Model}

\subsubsection{System Model}
\begin{itemize}
	\item Coevolution
	\item Self-organization
	\item Emergence	
	\item Adaption
\end{itemize}

Coevolution consist of:
\begin{itemize}
	\item Scarcity of customers
	\item Conscious choice that enables the organizations to change
	\item Interconnectedness of the enables the organizations to change
	\item Feedback processes that carry the long-term consequences of coevolution
\end{itemize}

\subsubsection{Management Model}

% subsection modeling (end)

\subsection{Mathematical Formulation} % (fold)
\label{sub:mathematical_formulation}
\subsubsection{Comprehended Model}
% subsection mathematical_formulation (end)

\subsection{Model Validation} % (fold)
\label{sub:model_validation}

% subsection model_validation (end)

% section methodological_approach (end)