### 2016-06-02 method.tex
添加新条目

% !TEX root = Eco-Model.tex
\section{Methodological approach} % (fold)
\label{sec:methodological_approach}

This section starts with a brief overview of ABMS. Then .... Finally ... 
i.e. the swarm model of the system.

In a word, the center of the research is the design of the user agent, and the core of the agent is to make decisions. so the core of the evolution is to make decision?

core of the paper: agent simulation, decision making.

\subsection{Framework of Agent-based modeling} % (fold)
\label{sub:framework of agent based modeling}
\begin{figure}[htbp]
\centering\small
\subfloat[Order Flow]{\resizebox{0.45\textwidth}{!}{\input{figures/order_flow}}\label{fig:orderflow}} \hspace{0.09\textwidth}
\subfloat[User Rank]{\resizebox{0.45\textwidth}{!}{\input{figures/user_rank}}\label{fig:userrank}}
% \resizebox{0.9\textwidth}{!}{\input{figures/simple_order_flow}}
\caption{Simple Flow Chart of Order}
% \label{fig:simpleorderflow}
\end{figure}

\subsubsection{Assumptions and notation}

\subsection{Agents in the ecosystem model} % (fold)
\label{sub:agents_in_the_ecosystem_model}
Using UML 
% subsection agents_in_the_ecosystem_model (end)

\subsubsection{Meta agent} % (fold)
\label{ssub:meta_agent}
is a decision maker
% subsubsection meta_agent (end)

\subsubsection{User agent} % (fold)
\label{ssub:user_agent}

% subsubsection user_agent (end)

\subsubsection{Platform agent} % (fold)
\label{ssub:platform_agent}

% subsubsection platform_agent (end)

\subsection{Ecosystem model} % (fold)
\label{sub:ecosystem model}

% subsection model_realization (end)

\subsubsection{Main modules design}
\label{subsub:main_modules}
inherent relationships, execution sequence/queue.

for service provider to modify their service area, i.e. breadth and depth.

Decision making involves choosing some course of action among various alternatives. For a decision maker, fulfilling the con

\subsubsection{System Model}
\begin{itemize}
    \item Co-evolution
    \item Self-organization
    \item Emergence 
    \item Adaption
\end{itemize}

Assumptions:
\begin{enumerate}
    \item substitute transport with time consume
    \item manufacturing agent randomly show up in the map
\end{enumerate}

Co-evolution consist of:
\begin{itemize}
    \item Scarcity of customers
    \item Conscious choice that enables the organizations to change
    \item Interconnectedness of the enables the organizations to change
    \item Feedback processes that carry the long-term consequences of co-evolution
\end{itemize}

% subsection modeling (end)

\subsection{Simulation design} % (fold)
\label{sub:simulation_design}

% subsection simulation_design (end)

\subsection{Agent-based modeling and simulation} % (fold)
\label{sub:agent_based_modeling_and_simulation}
\subsubsection{Agent design}

\subsubsection{Logical process}

\subsubsection{Interaction among agents}
some mathematical formulation display here to describe the whole ecosystem wide phenomenon, like Little formulation.

% subsection agent_based_model_and_simulation (end)
% section methodological_approach (end)