### 2016-08-12 method.tex
start to structure the paper

% !TEX root = Eco-Model.tex
\section{Methodological approach} % (fold)
\label{sec:methodological_approach}

This section starts with a brief overview of ABMS. Then .... Finally ... 
i.e. the swarm model of the system.

In a word, the center of the research is the design of the user agent, and the core of the agent is to make decisions. so the core of the evolution is to make decision?

core of the paper: agent simulation, decision making.

\subsection{Framework of Agent-based modeling} % (fold)
\label{sub:framework of agent based modeling}
\begin{figure}[htbp]
\centering\small
\subfloat[Order Flow]{\resizebox{0.45\textwidth}{!}{\input{figures/order_flow}}\label{fig:orderflow}} \hspace{0.09\textwidth}
\subfloat[User Rank]{\resizebox{0.45\textwidth}{!}{\input{figures/user_rank}}\label{fig:userrank}}
% \resizebox{0.9\textwidth}{!}{\input{figures/simple_order_flow}}
\caption{Simple Flow Chart of Order}
% \label{fig:simpleorderflow}
\end{figure}

\subsubsection{Assumptions and notation}
\begin{itemize}
    \item one provider can publish only one order at one time;
    \item service quality give the initial rank, the product quality depends on the service quality, review depends on the service quality, and the rank depends on the review;
    \item when it comes to rank value, we refer two different rank values: user rank(provider) and service rank, their relationship should be determined.
\end{itemize}

\subsection{Agents in the ecosystem model} % (fold)
\label{sub:agents_in_the_ecosystem_model}
Using UML 
% subsection agents_in_the_ecosystem_model (end)

\subsubsection{Meta agent} % (fold)
\label{ssub:meta_agent}
is a decision maker
% subsubsection meta_agent (end)

\subsubsection{Demander agent} % (fold)
\label{ssub:demander_agent}
After pricing, chose the provider and the amount considering the time limit of the project and the budget.
% subsubsection demander_agent (end)

\subsubsection{Provider agent} % (fold)
\label{ssub:provider_agent}
Pricing function specify two main parts:
unit cost and
provided amount
% subsubsection provider_agent (end)

\subsubsection{Complex user agent} % (fold)
\label{ssub:complex_user_agent}

% subsubsection complex_user_agent (end)

\subsubsection{Platform agent} % (fold)
\label{ssub:platform_agent}
matching the 
\begin{figure}

\begin{tikzpicture}
\begin{class}[text width=8cm]{Job}{0,0}
    \attribute{needCapacity : Map}
    \operation{lala(haha): void}
\end{class}
\end{tikzpicture}
\caption{Test UML Diagram}

\end{figure}
% subsubsection platform_agent (end)

\subsection{Ecosystem model} % (fold)
\label{sub:ecosystem model}

% subsection model_realization (end)

\subsubsection{Main flow design} % (fold)
\label{ssub:main_flow_design}
\begin{itemize}
    \item generate provider and demander
    \item generate order (with task)
    \item generate resources
    \item process tasks (with only resource)
    \item organize service
    \item process task (plus service)
    \item Review and elimination of provider and service
\end{itemize}
% subsubsection main_flow_design (end)

\subsubsection{Main modules design}
\label{subsub:main_modules}
inherent relationships, execution sequence/queue.

for service provider to modify their service area, i.e. breadth and depth.

Decision making involves choosing some course of action among various alternatives. For a decision maker, fulfilling the con

\subsubsection{System Model}
\begin{itemize}
    \item Co-evolution
    \item Self-organization
    \item Emergence 
    \item Adaption
\end{itemize}

Assumptions:
\begin{enumerate}
    \item substitute transport with time consume
    \item manufacturing agent randomly show up in the map
\end{enumerate}

Co-evolution consist of:
\begin{itemize}
    \item Scarcity of customers
    \item Conscious choice that enables the organizations to change
    \item Interconnectedness of the enables the organizations to change
    \item Feedback processes that carry the long-term consequences of co-evolution
\end{itemize}

% subsection modeling (end)

\subsection{Simulation design} % (fold)
\label{sub:simulation_design}

% subsection simulation_design (end)

\subsection{Agent-based modeling and simulation} % (fold)
\label{sub:agent_based_modeling_and_simulation}
\subsubsection{Agent design}

\subsubsection{Logical process}

\subsubsection{Interaction among agents}
some mathematical formulation display here to describe the whole ecosystem wide phenomenon, like Little formulation.

% subsection agent_based_model_and_simulation (end)
% section methodological_approach (end)