### 2016-08-20 figures/scheduleChart.tex
start to tell the story
assumptation, equation, module, review, intro, review module

% !TEX root = flow_head.tex

\begin{tikzpicture}
\foreach \x in {0,6,12,18}
	\draw [->,thick] (15.7cm,\x cm) node{t} (0,\x cm) -- +(15.5cm, 0) ;
\foreach \x in {0,6,12,18}
	\draw [->,thick] (0,\x cm) -- +(0,5.2 cm);
\foreach \x in {0,6,12,18}
	\draw[yshift=5.35cm] (0,\x cm) node{Capacity};
\foreach \x in {1,2,3,4}
	\draw[yshift=23.5cm] (7.5cm,-6*\x cm) node{$r_\x$};
\foreach \x in {1,...,15}
	\foreach \y in {0,6,12,18}
		\draw (\x cm,\y cm) -- +(0,2mm);
\foreach \x in {1,...,15}
	\foreach \y in {0,6,12,18}
		\draw[yshift=-2mm] (\x cm,\y cm) node{$\x$};
\foreach \y in {0,6,12,18}
	\foreach \x in {1,...,10}
		\draw[yshift=\y cm] (0,0.5*\x cm) -- +(2mm,0) (-3mm,0.5*\x cm) node{$\x$};

\fill[blue!20!,yshift=18cm] (5cm,4cm) rectangle (7cm,5cm);
\draw[yshift=18cm] (2cm,0) rectangle (5cm,3cm) (5cm,4cm) rectangle (8cm,5cm) (8cm,0) rectangle (13cm,3.5cm) (3.5cm,1.6cm) node{$t_3$} (6.5cm,4.5cm) node{$sc_2$} (10.5cm,1.8cm) node{$t_2$};
\draw[dotted,yshift=18cm] (0,5cm) -- (7cm,5cm) (0,4cm) -- (15cm,4cm) (7cm,0) -- (7cm,5cm);
\draw[yshift=12cm] (0,0) rectangle (7cm,2.5cm) (7cm,4cm) rectangle (8cm,5cm) (8cm,0) rectangle (13cm,2.5cm) (3.5cm,1.35cm) node{$t_1$} (7.5cm,4.5cm) node{$sc_2$} (10.5cm, 1.35cm) node{$t_2$};
\draw[dotted,yshift=12cm] (0,5cm) -- (7cm,5cm) (0,4cm) -- (15cm,4cm) (7cm,5cm) -- (7cm,0) (8cm,5cm) -- (8cm,0);
\draw[yshift=6cm] (0,0) rectangle (7cm,3.5cm) (7cm,2cm) rectangle (8cm,5cm) (3.5cm,1.85cm) node{$t_1$} (7.5cm,3.5cm) node{$sc_2$};
\draw[dotted,yshift=6cm] (0,5cm) -- (7cm,5cm) (8cm,2cm) -- (15cm,2cm) (8cm,0) -- (8cm,2cm);
\draw (0,3.5cm) rectangle (1cm,5cm) (2cm,0) rectangle (5cm,2.5cm) (8cm,0) rectangle (13cm,3.5cm) (3.5cm, 1.35cm) node {$t_3$} (0.5cm,4.3cm) node{$sc_1$} (10.5cm,1.8cm) node{$t_2$};
\draw[dotted] (0,3.5cm) -- (15cm,3.5cm) (1cm,0) -- (1cm,3.5cm);
\end{tikzpicture}