### 2016-08-30 head.tex

% !TEX root = Eco-Model.tex

\title{The Evolution of Manufacturing Ecosystem in Cloud Manufacturing Architecture
%Cloud Manufacturing Ecosystem - Scheduling And Evolving
}

\author[label1]{Shengkai Chen}
\author[label1]{Shuiliang Fang}
% \address[label1]{ZJU}
% \address[label1]{Address One}
\address[label1]{Zhejiang University, Institute of manufacturing engineering and automation, Hangzhou City, Zhejiang Province, China}

% \fntext[label4]{Small city}

\begin{abstract}
In Cloud Manufacturing, individuals engage in manufacturing business through the well-designed platform framework, where provider provide manufacturing resource for demander to search and purchase it as product. This business model allows centralized management of distributed massive manufacturing resources, and could potentially reduce the idle rate of those resources.
However, in the manufacturing ecosystem constructed, the heterogeneity of resource's attribute like type, quality, price, processing velocity, leads tough taxonomy of these resources, and the chaotic operating mode, makes it hard to meet demander's need. Meantime, in order to enhance the cash flow rate, provider make decisions to enhance the use rate of their own resources.
The relationship among participants in this ecosystem become more complicated than that in normal manufacturing system, since the integrated advanced technologies within cloud environment makes it possible to let one individual make decisions depend on the surroundings and other's decisions. Hence, in this paper, we designed an original operation mode with 3 extension options for the cloud manufacturing ecosystem to support some decision makings for individuals and the platform operator.
With the help of Repast Simphony, we designed an agent-based model and experiment to simulate these operation modes with any combinations to validate the effectiveness of them. As the experiment result, incubation mode can help reduce the waiting of job and accelerate the process velocity of resources, outsourcing mode reduce the job queue length but not as much as incubation mode and requires more resources, metabolism mode require lower resources than the other modes and will slow resource process velocity a little. The combination of incubation and metabolism mode is ideal for platform operators to maintain the manufacturing ecosystem.
\end{abstract}

\begin{keyword}
Cloud manufacturing ecosystem \sep
Resource servitization and optimization\sep
Operation mode\sep
Manufacturing evolution\sep
Agent-based simulation

\end{keyword}