### 2016-08-30 intro.tex

% !TEX root = Eco-Model.tex
\section{Introduction}
With the emergence of cloud manufacturing, the relationship among participants in this manufacturing system become more complicated than that in normal manufacturing system, since the integrated advanced technologies within cloud environment makes it possible to let one individual make decisions depend on the surroundings and other's decisions. Demander prefer high quality performance on their needs while provider prefer high use rate of their resource, even the cloud platform operator prefer lower amount of registered resource when dealing with almost arriving needs for the easier maintenance. These preferences stimulate the emergence of good service pattern that will lead to the evolution of manufacturing pattern in the cloud circumstance like an ecosystem. Hence, hence it's important to identify good operation mode that help the evolution.

In this paper, we designed an original operation mode for the cloud manufacturing ecosystem, then we proposed 3 extension modes named metabolism mode, incubation mode and outsourcing mode, that can be extended on the original mode with any combinations. Finally, we deigned an experiment to validate these synthetic operation modes in agent-based simulation method.