### 2016-09-06 discussion.tex
改版

% !TEX root = Eco-Model.tex
\section{Limitations, and Future Research} % (fold)
\label{sec:contributions_limitations_and_future_research}
We only proposed 3 extensions to combine and it may not fully describe the operation mode of cloud manufacturing system, it may also limit the evolution direction of the ecosystem, hence we will design more extensions. The assignment of service-call is oversimplified to prevent complex consequence, we will design new approaches to assign this job.
% section contributions_limitations_and_future_research (end)

\section{Acknowledgments} % (fold)
\label{sec:acknowledgments}
This work is supported by the National Natural Science Foundation of China (No. 71571161), and the National High Technology Research and Development Program of China (863 Program) (No. 2015AA042101). 
% section acknowledgments (end)