### 2016-09-06 intro.tex

% !TEX root = Eco-Model.tex
\section{Introduction}
The relationship among individuals in cloud manufacturing system become more complicated than that in normal manufacturing system, since the integrated advanced technologies makes it possible to let one individual make decisions depend on the surroundings and others'. Demander prefer high quality performance while provider prefer high use rate of their resource, the cloud platform operator prefer lower amount of registered resource when dealing with almost arriving needs for an easier maintenance. These preferences stimulate the emergence of good service pattern that will lead to the evolution of manufacturing pattern in the cloud manufacturing ecosystem. Hence, it's important to identify good operation mode that help the evolution.

In this paper, we designed an original operation mode to describe the basic decision makings in cloud manufacturing ecosystem, then we proposed 3 extensions named metabolism mode, incubation mode and outsourcing mode. Finally, we deigned an experiment to validate these synthetic operation modes in an agent-based simulation method.