### 2016-09-08 head.tex
modify the abstract

% !TEX root = Eco-Model.tex

\title{The Evolution of Manufacturing Ecosystem in Cloud Manufacturing Architecture
%Cloud Manufacturing Ecosystem - Scheduling And Evolving
}

\author[add1]{Shengkai Chen}
\author[add1,add2]{Shuiliang Fang\corref{cor1}}
% \ead{me_fangsl@zju.edu.cn}
% \address[label1]{ZJU}
% \address[label1]{Address One}
\address[add2]{The State Key Laboratory of Fluid Power Transmission and Control, College of Mechanical Engineering, Zhejiang University, Hangzhou, 310027, China}
\address[add1]{Key Laboratory of Advanced Manufacturing Technology of Zhejiang Province, College of Mechanical Engineering, Zhejiang University, Hangzhou , 310027, China}
% \author[a,b,*]{Third Author}

% \address[a]{First affiliation, Address, City and Postcode, Country}
% \address[b]{Second affiliation, Address, City and Postcode, Country}
% \fntext[label4]{Small city}

\cortext[cor1]{Corresponding author. Tel.: +86-135-8880-3980; Email: me\_fangsl@zju.edu.cn}

\begin{abstract}
In cloud manufacturing, individuals engage in manufacturing business through a well-designed platform, where the operator is able to manage distributed massive manufacturing resources. 
Cloud manufacturing ecosystem has more complicated relationship among individuals in it than that in normal manufacturing system, one individual can make decisions depend on the surroundings and others'. In this paper, we designed an original operation mode with 3 extensions named incubation, outsourcing and metabolism for the ecosystem. We designed an agent-based simulation model to validate the effectiveness of these modes. Experiment result shows evolution in ecosystem: with incubation mode, the job queue length is reduced and the resource idle rate is declined; with outsourcing mode, the job queue length is also reduced but not as much as that with incubation mode and more resource is required; with metabolism mode, lower resource is required than that with other modes and resource idle rate just rose up a little.

\end{abstract}

\begin{keyword}
Cloud manufacturing ecosystem \sep
Decision making\sep
Operation mode\sep
Ecosystem evolution\sep
Agent-based simulation

\end{keyword}
% \belowfrontmatterskip0pt