### 2016-09-11 figures/scheduleChart.tex

% !TEX root = flow_head.tex

\begin{tikzpicture}
\foreach \x in {0,6.5,13,19.5}
	\draw [->,thick] (15.7cm,\x cm) node{t} (0,\x cm) -- +(15.5cm, 0) ;
\foreach \x in {0,6.5,13,19.5}
	\draw [->,thick] (0,\x cm) -- +(0,5.2 cm);
\foreach \x in {0,6.5,13,19.5}
	\draw[yshift=5.35cm] (0,\x cm) node{Capacity};
\foreach \x in {1,...,15}
	\foreach \y in {0,6.5,13,19.5}
		\draw (\x cm,\y cm) -- +(0,2mm);
\foreach \x in {1,...,15}
	\foreach \y in {0,6.5,13,19.5}
		\draw[yshift=-2mm] (\x cm,\y cm) node{$\x$};
\foreach \y in {0,6.5,13,19.5}
	\foreach \x in {1,...,10}
		\draw[yshift=\y cm] (0,0.5*\x cm) -- +(2mm,0) (-3mm,0.5*\x cm) node{$\x$};

\draw[yshift=6.5cm] (7.5cm,-11mm) node{$mr_4$, $C_{4,0} = 10$, $P(mr_4) = 3$};
\draw[yshift=13cm] (7.5cm,-11mm) node{$mr_3$, $C_{3,0} = 10$, $P(mr_3) = 3$};
\draw[yshift=19.5cm] (7.5cm,-11mm) node{$mr_2$, $C_{2,0} = 10$, $P(mr_2) = 2$};
\draw[yshift=26cm] (7.5cm,-11mm) node{$mr_1$, $C_{1,0} = 10$, $P(mr_1) = 1$};

% \fill[blue!20!,yshift=19.5cm] (6cm,4cm) rectangle (7cm,5cm);
\draw[yshift=19.5cm] (3cm,0) rectangle (6cm,3cm) (8cm,0) rectangle (13cm,3.5cm) (1cm,0) rectangle (3cm,4.5cm) (3cm,3cm) rectangle (6cm,3.5cm) (0,4.5cm) rectangle (1cm,5cm)  (4.5cm,1.6cm) node{$job_3$} (10.5cm,1.8cm) node{$job_2$} (2cm,2.25cm) node{$job_4$} (4.5cm,3.25cm) node{$job_5$} (0.5cm,4.75cm) node{$job_6$};
\draw[dotted,yshift=19.5cm] (3cm,4.5cm) -- +(12cm,0);

\draw[yshift=13cm] (0,0) rectangle (7cm,2.5cm) (7cm,2.5cm) rectangle (8cm,5cm) (8cm,0) rectangle (13cm,2.5cm) (1cm,2.5cm) rectangle (3cm,4.5cm)  (3.5cm,1.35cm) node{$job_1$} (7.5cm,3.75cm) node{$job_7$} (10.5cm, 1.35cm) node{$job_2$}  (2cm,3.5cm) node{$job_4$};
\draw[dotted,yshift=13cm] (0,5cm) -- (7cm,5cm) (8cm,2.5cm) -- (15cm,2.5cm);

\draw[yshift=6.5cm] (0,0) rectangle (7cm,3.5cm) (7cm,3cm) rectangle (8cm,5cm) (3cm,3.5cm) rectangle (6cm,5cm)  (8cm,0) rectangle (13cm,2.5cm) (3.5cm,1.85cm) node{$job_1$} (7.5cm,4cm) node{$job_7$} (4.5cm,4.25cm) node{$job_5$} (10.5cm,1.25cm) node{$job_2$};
\draw[dotted,yshift=6.5cm] (0,5cm) -- (7cm,5cm) (8cm,3cm) -- (15cm,3cm) (8cm,3cm) -- (8cm,2.5cm);

\fill[blue!20!] (6cm,2cm) rectangle (7cm,3.5cm);
\draw (0,3.5cm) rectangle (1cm,5cm) (3cm,0) rectangle (6cm,2.5cm)  (1cm,0) rectangle (3cm,3cm) (6cm,2cm) rectangle (8cm,3.5cm) (4.5cm, 1.35cm) node {$job_3$} (0.5cm,4.3cm) node{$job_6$}  (2cm,1.5cm) node{$job_4$} (7cm,2.75cm) node{$job_7$};
\draw[dotted] (0,3.5cm) -- (6cm,3.5cm) (1cm,0) -- (1cm,3.5cm) (8cm,2cm) -- (15cm,2cm) (7cm,2cm) -- +(0,1.5cm) (8cm,2cm) -- (8cm,0);
\end{tikzpicture}