### 2016-09-13 intro.tex

% !TEX root = Eco-Model.tex
\section{Introduction}
The relationship among entities in a cloud manufacturing system become more complicated than that in existing manufacturing systems, since the integration of advanced technologies makes it possible for individual to make decisions depend on enriched information. Demanders prefer high quality of performance, while providers prefer high use rate of resource. A cloud platform operator prefers a suitable amount of registered resources to satisfy the needs of the market and for an easier life cycle management of resources. These preferences stimulate the emergence of good service pattern that will lead to the emergence operation mode in the cloud manufacturing ecosystem. Hence, it's important to identify a suitable operation mode to meet most of individuals' preference and to optimize the resource management.

In this paper, an original operation mode is designed to describe the basic decision-makings in cloud manufacturing ecosystem, then three extensions, namely metabolism mode, incubation mode and outsourcing mode are proposed. Finally, an experiment to validate these synthetic operation modes are designed using an agent-based simulation method.