### 2016-11-03 review.tex

% !TEX root = Eco-Model.tex
\section{Background and literatures} % (fold)
\label{sec:literature_review}
\subsection{Background} % (fold)
\label{sub:background}
Resource consumption in manufacturing activities is inevitable, waste and idle of these resources  are pervasive in existing manufacturing systems \cite{Smith2012}. Effective utilization of resources and resource productivity driven manufacturing innovation is one key consideration to handle the environmental burden \cite{Dornfeld2014,Li2012}.

In cloud manufacturing context, provider entity publishes manufacturing resources, demander entity publishes manufacturing tasks, platform arranges these resources properly to perform the tasks.
In scope are the mode to generate manufacturing service, which is an arrangement of resources, and the mode to keep resource in high quality instead of low quality.

% subsection background (end)

\subsection{Literatures} % (fold)
\label{sub:literatures}

Platform operator can manage manufacturing service that encapsulated distributed manufacturing resources intensively with appropriate business model \cite{Xu2012}, modular approaches and multi-layer architectures are the most common approaches to build a cloud manufacturing platform or system framework \cite{Tao2012,Valilai2013}, Lv used the list of views to depict this multi-layer architecture \cite{LvJuly312012-Aug.22012}.

Servitization is the key philosophy to operate cloud manufacturing \cite{li2010cloud}. A  service can be created statically which comes along with a provider \cite{Tao2012}, or can be created dynamically according to task pattern, such method as `Multi-Composition for Each Task' \cite{Liu2013} that combines incompetent service as a whole. A service can also be created by AI planning-based automatic composition framework \cite{OhJan.-March2008}.

Simulation approach has been widely used in manufacturing systems on operation planning and scheduling, real-time control, operating policies, performance analysis \cite{Smith2003}. In operating policies field, scheduling policies can be tested with simulation performance under given machine conditions \cite{Sabuncuoglu2003}, machine
segmentation policies can be simulated in a combined MRP and Kanban production system \cite{Felberbauer9-12Dec.2012}. Mourtzis et al. \cite{Mourtzis2015} explored a series of simulation-based solutions in industrial practices and concluded that research trends are in  Internet- and cloud-based situations.

% subsection literatures (end)
% section literature_review (end)